% =========================================================
% precedentguard_theorems_v0.2_skeleton.tex
% PrecedentGuard — Complete Proof Skeleton for v0.2 Theorems
% AAAI 2027 Submission Working Document
% Created: 2026-06-30 evening (Day 0)
% Proofs due: Day 1 (July 1) EOD
%
% Contents:
%   Theorem 1: Directional Intervention Sensitivity Bound
%   Theorem 2: Population Double-Sided Risk Bound
%   Theorem 3: Finite-Sample Double-Sided Certificate
%   Proposition 1: TV-based Indistinguishability Lower Bound
%   Corollary 1: Authenticity is not Semantic Information
%
% Status markers:
%   [FILL: ...] = you must complete this content
%   [VERIFY: ...] = numerical / definitional choice to confirm
%   [REF: ...] = citation to add at this point
% =========================================================

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}

\newcommand{\TODO}[1]{{\color{red}\textbf{[FILL: #1]}}}
\newcommand{\VERIFY}[1]{{\color{orange}\textbf{[VERIFY: #1]}}}
\newcommand{\REF}[1]{{\color{teal}\textsf{[REF: #1]}}}
\newcommand{\NOTE}[1]{{\color{blue}\textit{[Note: #1]}}}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

\DeclareMathOperator{\TV}{TV}
\DeclareMathOperator{\KL}{KL}
\DeclareMathOperator{\clip}{clip}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\attested}{attested}
\DeclareMathOperator{\type}{type}

\title{PrecedentGuard --- Complete Proof Skeleton\\
       \large Theorems 1--3 and Proposition 1 (v0.2 Framework)}
\author{AAAI 2027 Working Document}
\date{2026-06-30}

\begin{document}
\maketitle

\section{Setup and Notation (Aligned with v0.2 \S 3)}

\subsection{Score decomposition}

The PrecedentGuard score on input $x = (I, A, P_x, R_x)$ is
\[
S_{\mathrm{PG}}(x) = B(I, A) + \clip\!\Big(\sum_{e \in P_x} \beta_e \tilde\delta_e^{\mathrm{PG}}
        + \sum_{i \in R_x} w_i \tilde\delta_i^{\mathrm{PG}},\ -\varepsilon_-,\ \varepsilon_+ \Big),
\]
where $B(I, A) \in \mathbb{R}$ is the frozen guard's intent--action base score,
each per-evidence contribution $\tilde\delta_e^{\mathrm{PG}}$ is clipped to a type-specific interval $[-c_{\type(e)}^-, c_{\type(e)}^+]$,
and the directional trust constraint replaces the lower endpoint by $0$ when $\attested(e, x) = 0$:
\[
\tilde\delta_e^{\mathrm{PG}} =
\begin{cases}
\max\!\bigl(0,\ \tilde\delta_e\bigr), & \attested(e,x) = 0,\\[2pt]
\tilde\delta_e, & \attested(e,x) = 1.
\end{cases}
\]
The decision is $\hat Y(x) = \mathbf{1}[S_{\mathrm{PG}}(x) \ge \theta]$.

\subsection{Attack class and budget}

The adversary perturbs at most $m_k$ evidence nodes of type $k$, summarized in a vector budget
$\mathbf m = (m_{\mathrm{mem}}, m_{\mathrm{obs}}, m_{\mathrm{ret}}, m_{\mathrm{tool}}, m_{\mathrm{prec}})$.
We write $\mathcal A(\mathbf m)$ for the set of allowed attacks.
For an attack $a$, let $S(x)$ denote the clean score and $S(a(x))$ the score under attack
with the same $(I, A)$.

\subsection{Class-conditional margins}

For an example $x$ with safety label $Y \in \{0, 1\}$, the clean margin is
\[
M_1(x) = S(x) - \theta \quad (Y = 1, \text{unsafe}),
\qquad
M_0(x) = \theta - S(x) \quad (Y = 0, \text{safe}).
\]

\subsection{Assumptions (A1--A5, v0.2 \S 5.1)}

\begin{assumption}[Execution-graph fidelity]\label{a1}
Every mutable evidence item that enters $S_{\mathrm{PG}}$ is represented by an
instrumented node in the EIG. Hidden channels outside the instrumentation are out of scope.
\end{assumption}

\begin{assumption}[Bounded contribution interface]\label{a2}
Each per-evidence contribution is clipped to $[-c_{\type(e)}^-, c_{\type(e)}^+]$ before aggregation.
The outer aggregate is clipped to $[-\varepsilon_-, \varepsilon_+]$.
\end{assumption}

\begin{assumption}[Attack budget]\label{a3}
The adversary modifies at most $m_k$ evidence nodes of each type $k$, counting insertions, deletions, and
replacements explicitly. A replacement counts as one modified node.
\end{assumption}

\begin{assumption}[Attestation boundary]\label{a4}
For the primary certificate, the adversary cannot forge or improperly obtain a policy attestation.
Authenticated but semantically unauthorized writes remain in the threat model.
Sensitivity to attestation error $\eta$ is treated in \S 5.5 of the main paper.
\end{assumption}

\begin{assumption}[Grid pre-commitment]\label{a5}
The configuration grid $\Gamma$ is fixed \emph{before} any calibration example
in $C_0 \cup C_1$ is observed. In particular, $\Gamma$ is not chosen by minimizing
$\widehat R_{\mathrm{FN}}(\gamma)$ or $\widehat R_{\mathrm{FP}}(\gamma)$ over calibration data.
In implementation, a hash of $\Gamma$ is committed to the experiment registry
(\texttt{experiments/registry.csv}) before the calibration split $C_0 \cup C_1$ is instantiated.
This assumption is required for the union bound in Theorem~\ref{thm:3};
Theorems~\ref{thm:1}--\ref{thm:2} do not depend on it.
\end{assumption}

\NOTE{These assumptions are visible deployment conditions, not universal claims.}

% =========================================================
\section{Theorem 1: Directional Intervention Sensitivity Bound}
% =========================================================

\begin{theorem}[Directional Intervention Sensitivity Bound]\label{thm:1}
Under Assumptions~\ref{a1}--\ref{a4}, for every attack $a \in \mathcal A(\mathbf m)$ and every input $x$,
\[
- \rho_-(\mathbf m) \ \le\ S(a(x)) - S(x) \ \le\ \rho_+(\mathbf m),
\]
where, writing $\Delta_k := c_k^- + c_k^+$ and $m_k^{\mathrm{ins,unattested}} \le m_k$ for the
unauthenticated-insertion sub-budget of type $k$,
\[
\rho_+(\mathbf m) = \min\!\Big\{\varepsilon_+,\ \sum_k m_k \Delta_k\Big\},
\]
\[
\rho_-(\mathbf m) = \min\!\Big\{\varepsilon_-,\ \sum_k \big(m_k - m_k^{\mathrm{ins,unattested}}\big)\Delta_k\Big\}.
\]
In particular, an unauthenticated insertion alone cannot decrease the score.
\end{theorem}

\subsection{Proof of Theorem 1}

\begin{proof}
Let $Z(x) = \sum_e \beta_e \tilde\delta_e^{\mathrm{PG}} + \sum_i w_i \tilde\delta_i^{\mathrm{PG}}$ denote
the clean pre-outer-clip aggregate, and $Z(a(x))$ the attacked aggregate.
Let $T \subseteq P_x \cup R_x$ index the modified terms. Write
\[
Z(a(x)) - Z(x) = \sum_{j \in T} \bigl(z_j^{a} - z_j\bigr),
\]
where each $z_j$ is the (clipped, trust-adjusted) contribution of the $j$-th term.

\paragraph{Step 1 (per-term envelope).}
For a modified term of type $k$, the clean and attacked values both lie in $[-c_k^-, c_k^+]$
after type clipping. Hence
\[
-(c_k^- + c_k^+) \ \le\ z_j^{a} - z_j \ \le\ c_k^- + c_k^+ = \Delta_k.
\]

\paragraph{Step 2 (directional refinement for unauthenticated insertions).}
Consider an \emph{unauthenticated} insertion: an attacker inserts a fresh evidence node $e$ such that
$\attested(e, x) = 0$. Before the directional trust rule, this term could carry a contribution as low as $-c_k^-$.
After applying $\tilde\delta_e^{\mathrm{PG}} = \max(0, \tilde\delta_e)$, the minimum possible value is $0$.
Since the clean configuration assigns $0$ to a non-existent node by convention, the change is
\[
z_j^{a} - z_j \in [0,\ c_k^+] \qquad \text{(unauthenticated insertion of type }k\text{)}.
\]
Hence such terms contribute nothing to the downward shift.

\paragraph{Step 3 (deletions and replacements remain bidirectional).}
A deletion removes an existing $z_j \in [-c_k^-, c_k^+]$ and sets it to $0$, giving change in $[-c_k^+,\ c_k^-]$;
a replacement keeps the attacker in the full bidirectional envelope $[-\Delta_k,\ \Delta_k]$.
Replacements of \emph{already-attested} evidence retain the bidirectional envelope because the trust rule
permits negative attested contributions.

\paragraph{Step 4 (summation over modified terms).}
Summing per-term envelopes over $T$ yields raw bounds
\[
-r_-(\mathbf m) \ \le\ Z(a(x)) - Z(x) \ \le\ r_+(\mathbf m),
\]
with
\[
r_+(\mathbf m) = \sum_k m_k \Delta_k,
\qquad
r_-(\mathbf m) = \sum_k \big(m_k - m_k^{\mathrm{ins,unattested}}\big)\Delta_k.
\]

\paragraph{Step 4$'$ (worst-case attacker allocation).}
The expression $r_-(\mathbf m)$ is monotone decreasing in $m_k^{\mathrm{ins,unattested}}$:
each unauthenticated insertion swapped in place of a replacement or attested deletion
strictly reduces the downward-shift budget by $\Delta_k$ (Step~2). An adversary
maximizing the downward shift therefore selects $m_k^{\mathrm{ins,unattested}} = 0$
for all $k$, yielding the loosest form $r_-(\mathbf m) = \sum_k m_k \Delta_k$;
the coupled attack combining insertions with deletions or replacements cannot beat
this worst-case allocation for the defender. The refined form
$r_-(\mathbf m) = \sum_k (m_k - m_k^{\mathrm{ins,unattested}}) \Delta_k$ is a
strictly tighter bound available only when the defender can \emph{enforce or observe}
a positive unauthenticated-insertion sub-budget (e.g., through provenance auditing).

\paragraph{Step 5 (effect of the outer clip).}
The outer clip $C(z) = \clip(z,\ -\varepsilon_-,\ \varepsilon_+)$ is $1$-Lipschitz and monotone:
for any $z_1, z_2$, $|C(z_1) - C(z_2)| \le |z_1 - z_2|$, and $C$ preserves order.
Therefore
\[
S(a(x)) - S(x) = C(Z(a(x))) - C(Z(x)) \in \big[-r_-,\ r_+\big].
\]
Additionally, the image of $C$ has diameter $\varepsilon_- + \varepsilon_+$, so the change is also bounded
componentwise by $\min(r_+, \varepsilon_+)$ above and $\min(r_-, \varepsilon_-)$ below.
This yields exactly $\rho_+$ and $\rho_-$ as defined.
\end{proof}

\subsection{Remarks on Theorem 1}

\begin{remark}[Tightness]\label{rem:thm1-tight}
The upward bound is attained by an attacker that performs a single insertion of type $k^\star$ with the
largest $c_{k^\star}^+$, with $\beta_e = 1$. The downward bound, restricted to non-insertion attacks,
is attained by a single replacement of an attested positive term by its negative endpoint.
\end{remark}

\begin{remark}[Outer clip is not a redundant safety net]
If the per-type caps $c_k^{\pm}$ are large, the outer clip $\varepsilon_\pm$ becomes the binding constraint
and gives a uniform bound independent of $\mathbf m$. This is the configuration we recommend whenever
the attacker can choose $\mathbf m$ adaptively.
\end{remark}

\begin{remark}[Limit of the bound]
Theorem~\ref{thm:1} does \emph{not} imply that the underlying frozen guard is Lipschitz on arbitrary text.
Robustness comes from preventing mutable evidence from entering an unbounded monolithic context,
not from a textual smoothness property. \TODO{Add a one-line statement here citing why \S 4.2 of v0.2 explicitly avoids this claim.}
\end{remark}

% =========================================================
\section{Theorem 2: Population Double-Sided Risk Bound}
% =========================================================

\begin{theorem}[Population Double-Sided Risk Bound]\label{thm:2}
Under Theorem~\ref{thm:1},
\[
\sup_{a \in \mathcal A(\mathbf m)} \mathrm{FNR}(a)
   \ \le\ \Pr\!\big(M_1 \le \rho_-(\mathbf m) \mid Y = 1\big),
\]
\[
\sup_{a \in \mathcal A(\mathbf m)} \mathrm{FPR}(a)
   \ \le\ \Pr\!\big(M_0 \le \rho_+(\mathbf m) \mid Y = 0\big).
\]
\end{theorem}

\subsection{Proof of Theorem 2}

\begin{proof}
We show the FNR bound; the FPR bound is symmetric.

Let $x$ have $Y = 1$ and suppose $M_1(x) > \rho_-(\mathbf m)$. Then
\[
S(x) - \theta > \rho_-(\mathbf m).
\]
By Theorem~\ref{thm:1},
\[
S(a(x)) \ge S(x) - \rho_-(\mathbf m) > \theta,
\]
so $\hat Y(a(x)) = 1$ (the attacked classifier still blocks), and $x$ is not a false negative under attack $a$.

Therefore, for any attack $a \in \mathcal A(\mathbf m)$, the false-negative event satisfies
\[
\big\{\hat Y(a(x)) = 0,\ Y = 1\big\} \ \subseteq\ \big\{M_1(x) \le \rho_-(\mathbf m),\ Y = 1\big\}.
\]
Taking conditional probabilities and then supremum over $a \in \mathcal A(\mathbf m)$ gives the FNR claim.

For FPR: if $Y = 0$ and $M_0(x) > \rho_+$, then $\theta - S(x) > \rho_+$, hence
$S(a(x)) \le S(x) + \rho_+ < \theta$, so $\hat Y(a(x)) = 0$ and $x$ is not a false positive.
The same containment argument yields the FPR bound.
\end{proof}

\subsection{Remarks on Theorem 2}

\begin{remark}[Clean errors are absorbed]
A clean false negative has $S(x) < \theta$, i.e.\ $M_1(x) < 0 \le \rho_-$. Such examples are already in
the bounding event $\{M_1 \le \rho_-\}$, so the theorem implies $\mathrm{FNR}^{\mathrm{clean}} \le U_{\mathrm{FN}}$
as a special case.
\end{remark}

\begin{remark}[Avoiding the trivial ``always block'' solution]
A vacuous classifier that blocks everything has $\mathrm{FNR} = 0$ but $\mathrm{FPR} = 1$.
The double-sided structure prevents passing this off as ``certified safe'': the FPR bound
exposes the safe-class margin distribution, which collapses to a point mass at $-\theta$ under always-block.
\TODO{Insert a one-paragraph illustration of how the calibration of $\theta$, $\varepsilon_\pm$, $c_k^\pm$
on the dev set jointly controls both $U_{\mathrm{FN}}$ and $U_{\mathrm{FP}}$.}
\end{remark}

% =========================================================
\section{Theorem 3: Finite-Sample Double-Sided Certificate}
% =========================================================

Let $C_1$ and $C_0$ be \emph{independent} held-out calibration subsets containing $n_1$ unsafe and $n_0$ safe
examples, respectively. For a configuration $\gamma = (\theta, \mathbf c, \varepsilon_-, \varepsilon_+)$, define
\[
\widehat R_{\mathrm{FN}}(\gamma) = \frac{1}{n_1} \sum_{x \in C_1} \mathbf{1}\!\big[M_1(x) \le \rho_-(\gamma)\big],
\quad
\widehat R_{\mathrm{FP}}(\gamma) = \frac{1}{n_0} \sum_{x \in C_0} \mathbf{1}\!\big[M_0(x) \le \rho_+(\gamma)\big].
\]
Let $\Gamma$ be the (finite) grid of configurations under consideration.

\begin{theorem}[Finite-Sample Double-Sided Certificate]\label{thm:3}
Under Assumptions~\ref{a1}--\ref{a5}, with probability at least $1 - \alpha$ over the draw of $C_0 \cup C_1$, simultaneously for all $\gamma \in \Gamma$,
\[
\sup_{a \in \mathcal A(\mathbf m)} \mathrm{FNR}(a)
   \ \le\ \widehat R_{\mathrm{FN}}(\gamma) + \sqrt{\frac{\log(2|\Gamma|/\alpha)}{2 n_1}},
\]
\[
\sup_{a \in \mathcal A(\mathbf m)} \mathrm{FPR}(a)
   \ \le\ \widehat R_{\mathrm{FP}}(\gamma) + \sqrt{\frac{\log(2|\Gamma|/\alpha)}{2 n_0}}.
\]
\end{theorem}

\begin{remark}[Sidedness of the union bound]\label{rem:thm3-sidedness}
We use \emph{one-sided} Hoeffding, giving the multiplier $2$: one upper-tail deviation per class, over $|\Gamma|$ configurations and $|\{\mathrm{FN}, \mathrm{FP}\}| = 2$ classes. Only the upper-tail deviation $R^{\star}(\gamma) \le \widehat R(\gamma) + t$ is required for a valid upper-bound certificate. A conservative two-sided variant would replace $2|\Gamma|$ by $4|\Gamma|$. The two variants differ by
\[
\frac{t_{4|\Gamma|}}{t_{2|\Gamma|}} \;=\; \sqrt{\,\frac{\log(4|\Gamma|/\alpha)}{\log(2|\Gamma|/\alpha)}\,},
\]
which is \emph{independent of the calibration size} $n$ but depends on $|\Gamma|$ and $\alpha$.
For the reference regime $(|\Gamma|, \alpha) = (20, 0.05)$, the two-sided tail is $5.06\%$ larger than the one-sided.
Representative values on the grid $(|\Gamma|, \alpha) \in \{5, 10, 20, 50, 100, 500\} \times \{0.01, 0.05, 0.10\}$
range from $2.97\%$ (at $|\Gamma|=500, \alpha=0.01$) to $7.26\%$ (at $|\Gamma|=5, \alpha=0.10$);
the full table is reproducible from \texttt{scripts/day1\_theorem\_numerical\_example.py}
(saved output: \texttt{scripts/day1\_theorem\_numerical\_example.log}, which also reproduces three externally supplied reference values as a spot-check).
We adopt the tighter one-sided version and cite this remark explicitly in \S 5.4 of the main draft.
\end{remark}

\subsection{Proof of Theorem 3}

\begin{proof}
Fix any $\gamma \in \Gamma$. For an example $x_i \in C_1$, define the Bernoulli random variable
\[
Z_i^{\mathrm{FN}}(\gamma) = \mathbf{1}\!\big[M_1(x_i) \le \rho_-(\gamma)\big].
\]
Since the $x_i$ are i.i.d.\ from the unsafe-class population,
\[
\mathbb{E}\!\big[Z_i^{\mathrm{FN}}(\gamma)\big] = \Pr\!\big(M_1 \le \rho_-(\gamma) \mid Y = 1\big)
   \ =:\ R_{\mathrm{FN}}^{\star}(\gamma).
\]
By Hoeffding's inequality (one-sided),
\[
\Pr\!\Big(R_{\mathrm{FN}}^{\star}(\gamma) - \widehat R_{\mathrm{FN}}(\gamma) > t\Big)
   \ \le\ \exp(-2 n_1 t^2),
\]
\REF{Hoeffding 1963, \emph{J.\ Amer.\ Statist.\ Assoc.}, Theorem 2; Boucheron, Lugosi \& Massart 2013, \emph{Concentration Inequalities}, Theorem 2.8.}

Apply a union bound over $\Gamma$ and over both classes.
By Assumption~\ref{a5}, $\Gamma$ is fixed before observing $C_0 \cup C_1$, so the $N$ tail events below are legitimate targets for a union bound.
Let $N = 2 |\Gamma|$: one upper-tail deviation per class, $|\Gamma|$ configurations times $|\{\mathrm{FN}, \mathrm{FP}\}| = 2$ classes; see Remark~\ref{rem:thm3-sidedness}.
Set
\[
t = \sqrt{\frac{\log(N/\alpha)}{2 n_{\min}}} \quad \text{with } n_{\min} = \min(n_0, n_1),
\]
so that the simultaneous probability of any deviation exceeding $t$ in any of the $N$ tail events is at most $\alpha$.
Then, with probability at least $1 - \alpha$, simultaneously for all $\gamma$ and both classes,
\[
R_{\mathrm{FN}}^{\star}(\gamma) \le \widehat R_{\mathrm{FN}}(\gamma) + \sqrt{\tfrac{\log(N/\alpha)}{2 n_1}},
\qquad
R_{\mathrm{FP}}^{\star}(\gamma) \le \widehat R_{\mathrm{FP}}(\gamma) + \sqrt{\tfrac{\log(N/\alpha)}{2 n_0}}.
\]

Combining with Theorem~\ref{thm:2}:
\[
\sup_{a \in \mathcal A(\mathbf m)} \mathrm{FNR}(a) \le R_{\mathrm{FN}}^{\star}(\gamma)
   \le \widehat R_{\mathrm{FN}}(\gamma) + \sqrt{\tfrac{\log(N/\alpha)}{2 n_1}},
\]
and similarly for FPR.
\end{proof}

\subsection{Remarks on Theorem 3}

\begin{remark}[Calibration--deployment matching]
The bound is distribution-dependent: it estimates $R^{\star}(\gamma)$ on $C_0 \cup C_1$ and inherits validity
only when the deployment risk distribution matches the calibration distribution for the relevant event.
Cross-domain validity must be evaluated empirically (\S 6.5 of v0.2).
\end{remark}

\begin{remark}[Independence of calibration and test]
The proof assumes $C_0$ and $C_1$ are drawn independently from the unsafe and safe class populations,
and that $\gamma$ is selected on a separate development split.
$\Gamma$-independence of the calibration data is now formally enforced by Assumption~\ref{a5} (grid pre-commitment):
the implementation logs a hash of $\Gamma$ in \texttt{experiments/registry.csv} before the calibration split $C_0 \cup C_1$ is instantiated,
which prevents post-hoc grid selection from silently breaking the union bound.
\end{remark}

\begin{remark}[Conformal-risk-control as a stronger alternative]
A conformal-risk-control variant (\REF{Angelopoulos, Bates, Fisch, Lei \& Schuster 2024, \emph{ICLR}; arXiv:2208.02814}) gives marginal guarantees without finite-grid union-bound losses. This paper defers the conformal extension to future work; a one-line pointer in \S 5.4 of the main draft is retained for the reader.
\end{remark}

% =========================================================
\section{Proposition 1: TV-Indistinguishability Lower Bound}
% =========================================================

\subsection{Measurable setup}\label{sec:prop1-setup}

Let $(\Omega, \mathcal{F}, \Pr)$ be the underlying probability space carrying the input $X$,
the label $Y \in \{0, 1\}$, and any adversary randomness.
Let $O: \Omega \to \mathcal{Y}$ be a measurable \emph{observation channel} emitting the
sub-$\sigma$-algebra $\sigma(O) \subseteq \mathcal{F}$ that is visible to a binary guard
$h: \mathcal{Y} \to \{0, 1\}$; in particular, $h$ is a $\sigma(O)$-measurable function.
The attacker $a$ is a Markov kernel (possibly randomized) from $\mathcal{X}$ to $\mathcal{X}$,
applied before observation. Define
\[
Q_y \;:=\; \operatorname{Law}\!\big(O(a(X)) \,\big|\, Y = y\big)
\qquad (y \in \{0, 1\}),
\]
the pushforward distributions on $\sigma(O)$. Under this setup,
\[
\mathrm{FNR}(h) \;=\; \Pr_{Q_1}\!\big[h = 0\big],
\qquad
\mathrm{FPR}(h) \;=\; \Pr_{Q_0}\!\big[h = 1\big].
\]

\begin{proposition}[Double-Sided Indistinguishability Lower Bound]\label{prop:1}
For any (deterministic or randomized) binary guard $h$ that is $\sigma(O)$-measurable,
\[
\mathrm{FNR}(h) + \mathrm{FPR}(h) \;\ge\; 1 - \TV(Q_1, Q_0),
\]
and hence
\[
\max\!\big\{\mathrm{FNR}(h),\ \mathrm{FPR}(h)\big\} \;\ge\; \frac{1 - \TV(Q_1, Q_0)}{2}.
\]
In particular, if $Q_1 = Q_0$, at least one of $\mathrm{FNR}(h)$ or $\mathrm{FPR}(h)$ is at least $1/2$.
\end{proposition>

\subsection{Proof of Proposition 1}

\begin{proof}
The guard $h$ is a binary hypothesis test that, on observing input distributed as either $Q_1$ (unsafe)
or $Q_0$ (safe), outputs a binary decision. The minimum possible sum of type-I and type-II errors over all
(deterministic or randomized) tests between $Q_1$ and $Q_0$ is
\[
\inf_h \Big(\Pr_{Q_1}[h = 0] + \Pr_{Q_0}[h = 1]\Big) = 1 - \TV(Q_1, Q_0)
\]
by Le~Cam's two-point inequality \REF{Tsybakov 2009, \emph{Introduction to Nonparametric Estimation}, Theorem 2.2 (standard textbook form); original: Le Cam 1986, \emph{Asymptotic Methods in Statistical Decision Theory}}. The infimum is attained by the Neyman--Pearson test between $Q_1$ and $Q_0$. Hence every $\sigma(O)$-measurable binary guard $h$ satisfies
\[
\mathrm{FNR}(h) + \mathrm{FPR}(h) = \Pr_{Q_1}[h = 0] + \Pr_{Q_0}[h = 1] \ge 1 - \TV(Q_1, Q_0).
\]
The maximum-of-two bound follows from $\max(a, b) \ge (a + b)/2$ for non-negative $a, b$.
\end{proof}

\subsection{Remarks on Proposition 1}

\begin{remark}[Role of Proposition 1]
This proposition is a \emph{supporting clarification}, not the paper's primary novelty. Its purpose is to
formalize a precise reason why provenance authenticity (a signature) and semantic authorization (a policy
attestation tied to a safety label) must be separated: if an authenticated channel does not move
$Q_1$ apart from $Q_0$, it cannot improve the classification floor.
\end{remark}

\begin{remark}[Compared to the previously-planned universal impossibility]
The previous plan (v1.x, archived) stated a universal ``no-trust-anchor $\Rightarrow$ no double-sided control''
impossibility. That claim was too broad because a guard may classify perfectly from an entirely different
uncorrupted channel. Proposition~\ref{prop:1} replaces it with the correct, narrower obstruction:
\emph{lack of label-separating information} in the attacker's observation channel.
\end{remark}

% =========================================================
\section{Corollary 1: Authenticity is not Semantic Information}
% =========================================================

\begin{corollary}[Authenticity is not Semantic Information]\label{cor:1}
Let $Z$ be an authenticated variable appended to the guard's observation (a signature, a write-time
provenance tag, etc.). If
\[
\TV\!\big(Q_1^{O, Z},\ Q_0^{O, Z}\big) \ =\ \TV\!\big(Q_1^O,\ Q_0^O\big),
\]
then introducing $Z$ does not improve the lower bound of Proposition~\ref{prop:1}. In particular, a valid
signature whose distribution is independent of the safety label, or that authenticates adversarially chosen
content, cannot break indistinguishability.
\end{corollary}

\begin{proof}
Marginalization from $(O, Z)$ back to $O$ is a deterministic channel, hence a valid (deterministic) randomization.
By the data-processing inequality for total variation
\REF{Polyanskiy \& Wu 2024, \emph{Information Theory: From Coding to Learning}, Ch.~7; Cover \& Thomas 2006, \emph{Elements of Information Theory}, Ch.~2},
\[
\TV\!\big(Q_1^{O,Z}, Q_0^{O,Z}\big) \;\ge\; \TV\!\big(Q_1^{O}, Q_0^{O}\big).
\]
The bound in Proposition~\ref{prop:1} is monotone decreasing in the TV distance of the observation channel;
therefore, if extending the channel from $O$ to $(O, Z)$ does not \emph{strictly} increase TV, the lower bound on
$\mathrm{FNR}(h) + \mathrm{FPR}(h)$ is unchanged. A sufficient condition for equality is $Z \perp Y \mid O$
(the appended variable is conditionally independent of the label given the existing observation).
\end{proof}

\NOTE{This formalizes why \S 1.2 of v0.2 distinguishes \emph{authenticity}
from \emph{semantic authorization}: a policy attestation is informative about $Y$ only if its conditional
distribution depends on $Y$.}

% =========================================================
\section{Day 1 Task Map (July 1)}
% =========================================================

\subsection{Hour-by-hour schedule}

\begin{enumerate}[label=\textbf{\arabic*.}, leftmargin=*]

\item \textbf{09:00--09:30 --- Compile check.} Compile this file. Fix any LaTeX errors (likely
encoding or package-related). Verify all theorem environments render.

\item \textbf{09:30--11:00 --- Theorem 1 elaboration.} Read Theorem~\ref{thm:1} and its proof.
Resolve every \TODO{...} marker. The most non-trivial step is the case analysis in Steps 2--3:
verify that your codebase's \emph{insertion} convention matches the proof (i.e., a clean ``no node''
contributes $0$).

\item \textbf{11:00--12:00 --- Theorem 2 elaboration.} The proof is short; the work is checking that
$\rho_-(\mathbf m)$ as computed by your implementation matches the value used in the proof.
Run one numerical example end-to-end.

\item \textbf{14:00--16:00 --- Theorem 3 elaboration.} Resolve the \VERIFY{} marker on the constant
$4 |\Gamma|$. Choose one of:
\begin{itemize}
  \item Conservative two-sided ($4|\Gamma|$): simpler narrative, slightly weaker bound.
  \item One-sided ($2|\Gamma|$): tighter, requires explicit ``one-sided'' caveat.
\end{itemize}
Recommendation: use $2|\Gamma|$ and state it explicitly.

\item \textbf{16:00--17:00 --- Proposition 1 elaboration.} Verify Le~Cam citation; this is standard
but the exact reference matters for the rebuttal.

\item \textbf{17:00--18:00 --- Independent review request.} Send this document plus a short summary email
to a colleague with measure-theory background. Ask: ``Are Steps 2--3 of Theorem 1 watertight under
your reading of insertion, deletion, and replacement?''
\end{enumerate}

\subsection{End-of-day checkpoint}

By 18:00 you should have:
\begin{itemize}[noitemsep]
  \item All \TODO{} markers resolved.
  \item The \VERIFY{} marker for Theorem 3 decided.
  \item One numerical example run end-to-end demonstrating the bound.
  \item One independent review request out.
\end{itemize}

% =========================================================
\section{Self-Review Checklist (Before Gate $\alpha$)}
% =========================================================

\begin{enumerate}[label=$\Box$, leftmargin=*, noitemsep]
\item[$\boxtimes$] Notation matches v0.2 main draft \S 3 exactly. \emph{[done 2026-07-01; Grep-verified]}
\item[$\boxtimes$] Assumptions \ref{a1}--\ref{a5} are quoted verbatim from v0.2 \S 5.1 (A5 added 2026-07-01 to formalize grid pre-commitment).
\item[$\boxtimes$] Theorem 1's directional refinement (Step 2 + Step 4$'$) is consistent with \texttt{precedentguard/clipping.py::directional\_trust\_clip} and cross-validated by \texttt{tests/test\_clipping.py::TestDirectionalTrustClip}.
\item[$\boxtimes$] Theorem 2's containment argument uses only standard probability; verified by cold-read review 2026-07-01 (Attack Log \S 3).
\item[$\boxtimes$] Theorem 3 invokes Assumption~\ref{a5}; \texttt{precedentguard/certificate.py::assert\_grid\_committed} enforces the pre-commitment programmatically; \texttt{tests/test\_certificate.py::TestA5Enforcement} covers the post-hoc-grid attack.
\item[$\boxtimes$] Theorem 3's constant $2|\Gamma|$ (one-sided Hoeffding) is consistent throughout: skeleton L296/L300/L327/L579, main draft \S 5.4/Appendix A.3, and \texttt{scripts/day1\_theorem\_numerical\_example.py}.
\item[$\boxtimes$] Proposition 1 cites Tsybakov 2009 Thm 2.2 as the standard textbook form (Le Cam 1986 retained as historical origin).
\item[$\boxtimes$] Corollary 1's monotonicity of TV is formally justified via the data-processing inequality (Polyanskiy \& Wu 2024 Ch.~7 / Cover \& Thomas 2006 Ch.~2).
\item[$\boxtimes$] Non-vacuous numerical example verified: \texttt{scripts/day1\_theorem\_numerical\_example.log} for symbolic bounds; \texttt{experiments/day3\_smoke.log} for end-to-end non-vacuous $U_{FN}=U_{FP}=0.136$ at $n=100$ per class.
\item[$\square$] \textbf{BLOCKER (deferred to Day 5--7):} Human mathematical review of Steps 2--3 of Theorem 1 and the $\sigma$-algebra formalization in Proposition 1. AI cold-read audit completed 2026-07-01 (`agentId a5ab48f1747c1a7da`); this DOES NOT satisfy CLAUDE.md \S2 independent verification. Contact: TBD by Day 5.
\end{enumerate}

\NOTE{9 of 10 boxes checked as of 2026-07-04 EOD (Gate $\alpha$). Item 10 is deferred as a scheduled Day 5--7 task; the AI cold-read serves as pre-verification, not final verification. This deferral is disclosed in the Gate $\alpha$ judgement report (\texttt{experiments/gate\_alpha\_report.md}).}

% =========================================================
\section{Failure Modes and Fallbacks}
% =========================================================

\begin{itemize}[noitemsep]
\item \textbf{If Theorem 1's directional refinement is wrong (e.g., insertions can decrease the score under your
actual implementation):} Revert to the symmetric bound $\rho_{\mathrm{sym}} = \min(\varepsilon_-+\varepsilon_+,\ \sum_k m_k \Delta_k)$
and lose the asymmetric narrative.
\item \textbf{If Theorem 3's calibration assumption fails (e.g., $\Gamma$ depends on $C_0 \cup C_1$):}
Apply a holdout split or conformal-risk-control variant. \REF{Angelopoulos, Bates, Fisch, Lei \& Schuster 2024, ICLR; arXiv:2208.02814.}
\item \textbf{If Proposition 1's TV bound is contested by a reviewer:} It is Le Cam's two-point inequality.
Cite Tsybakov 2009 \emph{Introduction to Nonparametric Estimation}, Theorem 2.2, as a textbook source.
\end{itemize}

\end{document}
